### 2016-06-17 Draft.tex
Edit through Sim2

\newcommand{\pdet}{p^{(det)}}

@@

\jarad{Need a roadmap. Provide section number in the paragraph below.}

In this study, we construct a time-to-detection N-mixture model for estimating abundance from removal-sampled point-count surveys.  
We explore model behavior through a series of simulations testing: (i) how well constant-detection models fit non-constant detection data and vice versa, (ii) the utility of finite-mixtures to model detection heterogeneity involving a non-constant detection subgroup, and (iii) model performance with covariates and random effects affecting both abundance and detection.  
Finally, we apply our model to Ovenbird data from the Minnesota Forest Breeding Bird Project.

@@

\section{Time-to-detection N-mixture models}

@@

Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$. 
If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$. Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.

A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e. $T_{sb}\stackrel{ind}{\sim} \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \lambda$.
Choosing another TTDD can allow for a systematic non-constant detection regime. 
For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
For consistency that will be clear when including explanatory variables, we use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.

With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
In these models, estimates of $N_s$ are unstable unless $\pdet>0.5$.
One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} \Po(\lambda)$. 
We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \Po\left(\lambda(1-\pdet)\right)$.  
Although alternative distributions could be considered, e.g. negative binomial, our experience with avian point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here. 
\jarad{Adam: add refs.}

@@

Due to the harried process of avian point counts, times to first detection are typically not recorded exactly, but are instead censored into $I$ intervals. 
Let $C_i$ for $i=1,\dots,I$ indicate the right endpoint of the $i$th interval then $C_I$ is the total survey duration and, letting $C_0=0$, the $i$th interval is $(C_{i-1},C_{i}]$. 
Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
Assuming independence amongst individuals and sites, we have $\vn_{s} \sim \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.

@@

To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ via a mixed effects linear regression with a log link, i.e. $\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.

To incorporate explanatory variables on detection probability, we let the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \overset{ind}{\sim} N(0,\\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.

@@

For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicity. 
\begin{subequations}
\begin{eqnarray} \label{eq:model}
n_s^{(obs)} &\sim& Po(\lambda_s) \\
\vn_s &\sim& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
p_{s1} &=& \gamma + F_{T|det}(C_1|\alpha,\varphi_s),\, p_{si} = F_{T|det}(C_i|\alpha,\varphi_s)-F_{T|det}(C_{i-1}|\alpha,\varphi_s)\\
\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2) \\
\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \overset{ind}{\sim} N(0,\tau_{D[j]}^2)
\end{eqnarray}
\end{subequations}

\jarad{May need to remove subnumbers, but I'll wait to see if we need them.}

We adopt a Bayesian approach and therefore require a prior over the model parameters.
To ease construction of a default prior for this model, we standardize all explanatory variables and then construct priors to be diffuse within a reasonable range of values.
Prior mean and standard deviation (sd) for the abundance intercept was set at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
Prior mean and sd for the detection intercept were chosen so that, based on an intercept-only non-mixture model with $\alpha=1$: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.  
Normal priors for fixed effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
All standard deviations and $\alpha$ were given half-Cauchy priors with location 0 and scale 1 for the untruncated Cauchy and Unif(0,1) prior on the mixture parameter $\gamma$ in the mixture models.
All scalar parameters were assumed independent \emph{a priori}.

@@

@@

@@

\section{Simulation studies}

@@

\subsection{Sim1 Results: mixture vs. non-mixture TTDDs}\label{sec:mixture}

@@

\subsection{Sim2: constant vs. non-constant detection mixture TTDDs}\label{sec:family}

In Sim2, we simulated intercept-only datasets from the 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families (Table \ref{tbl:sim2}).  
Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.  
However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.  
Differences among non-constant TTDDs were less pronounced.  
Gamma models were more accurate than others for nonpeaked mixture datasets.  
Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.

Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.

In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.  For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.

\begin{table}[ht]
\footnotesize\centering
\begin{tabular}{l|l|l|ccccc|ccccc}
 \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
 \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
  \hline
\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\ 
& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
\cline{2-13}
& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
\cline{2-13}
& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
   \hline
\end{tabular}
\vspace{0.5cm}\\
\begin{tabular}{l|l|l|ccccc|ccccc}
 \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
 \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
  \hline
\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\ 
& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
\cline{2-13}
& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
\cline{2-13}
& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
   \hline
\end{tabular}
\caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim2).  
Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
\end{table}

@@

In Sim3, we generated a single replicate of datasets from models including covariates and random effects for both detection and abundance.  
Results echoed those of the first two simulation studies, suggesting that the addition of covariates and random effects did not alter basic model behavior.
These results included the detection bias of non-mixture models used to fit mixture datasets (Figure \ref{pdet_cater_correct}), detection bias of exponential mixture models used to fit data mixtures with time-varying $F_{T1}$ components (Figure \ref{pdet_cater_family}), and more precise and accurate performance of exponential mixture models when data were from exponential mixtures.

@@

\section{Discussion}

>>>>

\newcommand{\pdet}{p^{(det)}}
\newcommand{\ind}{\stackrel{ind}{\sim}}

@@

Section \label{sec:data} provides a description of the interval-censored time-to-detection avian point count data under consideration.
Section \label{sec:model} introduces an N-mixture model for estimating abundance from removal-sampled point-count surveys and variety of TTDDs. In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
Section \label{sec:sim} provides three simulation studies to assess the impact of TTDD choice on detection probability. 
Section \label{sec:ovenbirds} analyzes an Ovenbird data set under the different TTDD to determine the impact this choice has on detection probability and therefore the estimate of abundance.
Finaly, section \label{sec:discuss} discusses

@@

\section{Time-to-detection N-mixture models} \label{sec:model}

@@

Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$. 
If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$. Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \ind f_{T|det}(t_{sb})$
where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.

A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e. $T_{sb}\ind \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \lambda$.
Choosing another TTDD can allow for a systematic non-constant detection regime. 
For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
We use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.
The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$.

With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
In these models, estimates of $N_s$ are unstable unless $\pdet>0.5$.
One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$. 
With this assumption, we can decompose $N_{s}$ into observed and unobserved portions: $n^{(obs)} \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
Although alternative distributions could be considered, e.g. negative binomial, our experience with avian point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here. 
\jarad{Adam: add refs.}

@@

Due to the harried process of avian point counts, times to first detection are typically not recorded exactly, but are instead censored into $I$ intervals. 
Let $C_i$ for $i=1,\dots,I$ indicate the right endpoint of the $i$th interval then $C_I$ is the total survey duration and, letting $C_0=0$, the $i$th interval is $(C_{i-1},C_{i}]$. 
Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.

@@

To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ via a mixed effects linear regression with a log link, i.e. $\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \ind N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.

To incorporate explanatory variables on detection probability, we let the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.

@@

For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit.

\begin{eqnarray*} 
n_s^{(obs)} &\ind& Po(\lambda_s) \\
\vn_s &\ind& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
p_{s1} &=& \gamma + F_{T|det}(C_1|\alpha,\varphi_s),\, p_{si} = F_{T|det}(C_i|\alpha,\varphi_s)-F_{T|det}(C_{i-1}|\alpha,\varphi_s)\\
\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \ind N(0,\sigma_{A[j]}^2) \\
\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \ind N(0,\tau_{D[j]}^2)
\end{eqnarray*}
We adopt a Bayesian approach and therefore require a prior over the model parameters.
To ease construction of a default prior for this model, we standardize all explanatory variables and then construct priors to be diffuse within a reasonable range of values.
Prior mean and standard deviation (sd) for the abundance intercept was set at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
Prior mean and sd for the detection intercept were chosen so that, based on an intercept-only non-mixture model with $\alpha=1$: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.  
Normal priors for fixed effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
All standard deviations and $\alpha$ were given half-Cauchy priors with location 0 and scale 1 for the untruncated Cauchy and Unif(0,1) prior on the mixture parameter $\gamma$ in the mixture models.
All scalar parameters were assumed independent \emph{a priori}.

@@

\jarad{Adam: do we want posterior predictive pvalues at all?}

@@

\jarad{Adam: do we want to remove DIC?}

@@

\section{Simulation studies} \label{sec:sim}

@@

\subsection{Mixture versus non-mixture TTDDs}\label{sec:mixture}

@@

\subsection{Constant vs. non-constant detection mixture TTDDs}\label{sec:family}

The previous section adressed model mis-specification in terms of the mixture component. 
Now we turn to model misspecification of the distribution family. 
We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using settings that are the same as in the previous section and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.

Table \ref{tbl:sim2} provides a summary of the ability of each model to capture the detection probability. 
Each quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak since the two parameters, exponential rate and mixing parameter, do not provide enough flexibility for the mixture exponential distribution to adequately model times to detection that have an initial increase and delayed mode. 
In this situation, the exponential model underestimates the actual detection probability.
In contrast, if there is no delayed peak, the exponential model typically overestimates the actual detection probability.

\begin{table}[ht]
\footnotesize\centering
\begin{tabular}{l|l|l|ccccc|ccccc}
 \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
 \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
  \hline
\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\ 
& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
\cline{2-13}
& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
\cline{2-13}
& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
   \hline
\end{tabular}
\vspace{0.5cm}\\
\begin{tabular}{l|l|l|ccccc|ccccc}
 \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
 \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
  \hline
\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\ 
& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
\cline{2-13}
& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
\cline{2-13}
& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
   \hline
\end{tabular}
\caption{Summary of mixture inference models of all families fit to mixture datasets.  
Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
\label{tbl:sim2}
\end{table}

When comparing the different three-parameter TTDDs, model misspecification is not as serious an issue because these models can adequately account for the interval-censored times to detection. 
Nonetheless, it appears the gamma mixture model is better able to account for data from lognormal or Weibull mixture models. 
In the gamma quadrant, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility. 
In contrast, when the lognormal model is used for inference, it performed worse when the data arose from a gamma or Weibull and, when the Weibull model was used for inference, it performed worse when the data arose from a lognormal.

Other than providing evidence against the exponential mixture model, DIC was not helpful in distinguishing amongst these models.

% Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.  
% However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.  
% Differences among non-constant TTDDs were less pronounced.  
% Gamma models were more accurate than others for nonpeaked mixture datasets.  
% Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
% Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
% 
% Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
% For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
% So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
% For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
% 
% In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.  For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.

@@

The previous sections studying the effect of the time to detection assumption in the context of no explanatory variables.
Here we incorporate fixed and random effects for abundance and detection. 
For each of the 14 combinations of TTDDs \jarad{Adam: what did we do? i don't remember how the data sets were constructed and therefore how to interpret the figures.}
We chose values for fixed effects and random effect variances similar to those found in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
Due to the computation time involved in estimating models with these fixed and random effects, we simulated two data sets: mixture and non-mixture.

To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.

Figure \ref{pdet_cater_correct}

These results included the detection bias of non-mixture models used to fit mixture datasets (Figure \ref{pdet_cater_correct}), detection bias of exponential mixture models used to fit data mixtures with time-varying $F_{T1}$ components (Figure \ref{pdet_cater_family}), and more precise and accurate performance of exponential mixture models when data were from exponential mixtures.

@@

\section{Discussion} \label{sec:discuss}
